Resolvable Ambiguity

### 2019-07-11 paper/paper.tex
Minor updates in section 4.

A formal language is defined as a set of words, i.e., a subset of $\T^{*}$ for some alphabet $\T$. For defining \emph{resolvable ambiguity} we additionally have to consider the results of parsing words. In order to stay as general as possible, we first define the notion of an \emph{abstract parser}.

@@

This notion of an abstract parser enables the introduction of a particular class of formal languages that we will use throughout the rest of the paper; we call members of this class \emph{parse languages:}

@@

\noindent For example, consider a simple arithmetic language without precedence and parentheses. In such a language, $\parse(1 + 2 \cdot 3)$ would produce a set containing the two parse trees shown in Fig.~\ref{fig:arith-example}.

\begin{figure}[t]
\begin{tabular}{cc}
  \Tree [.{$+$}
    1
    [.{$\cdot$}
      2
      3 ] ] &
  \Tree [.{$\cdot$}
    [.{$+$}
      1
      2 ]
    3 ] \\
\end{tabular}
\caption{The two allowable parse trees of $1 + 2 \cdot 3$ in an arithmetic language without precedence.}
\label{fig:arith-example}
\end{figure}

The ambiguity of a word $w \in \T^{*}$ is defined in terms of $\parse$:

\begin{definition}
  Given an abstract parser $P = (\T, T, \parse)$, a word $w \in L(P)$
  is ambiguous, written $\amb_P(w)$, iff

  $\exists t_1, t_2 \in T.\ t_1 \neq t_2 \land \{t_1, t_2\} \subseteq \parse(w)$
\end{definition}

@@

The example in Fig.~\ref{fig:arith-example} is not resolvably ambiguous, since neither tree can be written in any other way. However, if we add grouping parentheses to the language we find that $1 + (2 \cdot 3)$ and $(1 + 2) \cdot 3$ are unambiguously parsed as the first and second tree, respectively.

@@

Before introducing this formalism however, we introduce the two central problems we consider in this paper:

@@

\begin{definition}[Soundness of Static Resolvability]\label{def:static-procedure-sound}
  A decision procedure $f$ solving the static resolvability problem is sound iff

@@

\begin{definition}[Completeness of Static Resolvability]\label{def:static-procedure-complete}
  A decision procedure $f$ solving the static resolvability problem is \emph{complete} iff

@@

\begin{definition}[Soundness of Dynamic Resolvability]\label{def:dynamic-procedure-sound}
  A decision procedure $f$ solving the dynamic resolvability problem is sound iff

@@

Analogous to the static case, for the parse languages studied in the
following sections we can also provide a decision procedure $g$ such
that $g(\parse, w) \implies \lnot \rho_P(w) \text{ for all } w \in
\Sigma^{*} \text{ and abstract parsers } P = (\T, T, \parse)$.

\begin{definition}[Completeness of Dynamic Resolvability]\label{def:dynamic-procedure-complete}
  A decision procedure $f$ solving the dynamic resolvability problem is complete iff

>>>>

A formal language is defined as a set of words, i.e., a subset of $\T^{*}$ for some alphabet $\T$. To be able to define \emph{resolvable ambiguity}, we additionally have to consider the results of parsing words. In order to stay as general as possible, we first define the notion of an \emph{abstract parser}.

@@

\noindent This notion of an abstract parser enables the introduction of a particular class of formal languages that we will use throughout the rest of the paper; we call members of this class \emph{parse languages:}

@@

\noindent For example, consider a simple arithmetic language without precedence and parentheses. In such a language, $\parse(1 + 2 \cdot 3)$ would produce a set containing two parse trees. %the two parse trees shown in Fig.~\ref{fig:arith-example}.
%
%\begin{figure}[t]
%\begin{tabular}{cc}
%  \Tree [.{$+$}
%    1
%    [.{$\cdot$}
%      2
%      3 ] ] &
%  \Tree [.{$\cdot$}
%    [.{$+$}
%      1
%      2 ]
%    3 ] \\
%\end{tabular}
%\caption{The two allowable parse trees of $1 + 2 \cdot 3$ in an arithmetic language without precedence.}
%\label{fig:arith-example}
%\end{figure}
%
The ambiguity of a word $w \in \T^{*}$ is defined in terms of $\parse$:

\begin{definition}
  Given an abstract parser $P = (\T, T, \parse)$, a word $w \in L(P)$
  is ambiguous, written $\amb_P(w)$, iff
%
  $\exists t_1, t_2 \in T.\ t_1 \neq t_2 \land \{t_1, t_2\} \subseteq \parse(w)$
\end{definition}

@@

The example $\parse(1 + 2 \cdot 3)$
%in Fig.~\ref{fig:arith-example}
is not resolvably ambiguous, since neither of the two resulting trees can be written in any other way. However, if we add grouping parentheses to the language, we find that $1 + (2 \cdot 3)$ and $(1 + 2) \cdot 3$ are unambiguously parsed as the first and second tree, respectively.

@@

Before introducing this formalism, however, we introduce the two central problems we consider in this paper:

@@

\begin{definition}[Soundness of Static Resolvability]\label{def:static-procedure-sound}
  A decision procedure $f$, solving the static resolvability problem, is sound iff

@@

\begin{definition}[Completeness of Static Resolvability]\label{def:static-procedure-complete}
  A decision procedure $f$, solving the static resolvability problem, is \emph{complete} iff

@@

\begin{definition}[Soundness of Dynamic Resolvability]\label{def:dynamic-procedure-sound}
  A decision procedure $f$, solving the dynamic resolvability problem, is sound iff

@@

\noindent Analogous to the static case, for the parse languages studied in the
following sections, we can also provide a decision procedure $g$ such
that $g(\parse, w) \implies \lnot \rho_P(w) \text{ for all } w \in
\Sigma^{*} \text{ and abstract parsers } P = (\T, T, \parse)$.

\begin{definition}[Completeness of Dynamic Resolvability]\label{def:dynamic-procedure-complete}
  A decision procedure $f$, solving the dynamic resolvability problem, is complete iff
